---
title: "Choose_var_2017SP"
author: "TRS"
date: "11/23/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Choosing variable to run an ECOSPAT model with the Salt Pond 2017 data.

Import data table.

```{r}
inc_data_inc = read.table("species_for_varcorr.txt", header=T, row.names =1, sep ="\t")

```

Create Pearson pairwise correlation table. You want to avoid variables that are too correlated. This step may also help reduce the number of varibles from a long list.

```{r}
cor.df <- cor(inc_data_inc, method = "pearson")

```

Calculate the distances from the pairwise matrix.

```{r}
var.dist <- abs(as.dist(cor.df))
```

Run hierarchical clustering algorithm.

```{r}
var.cluster <- hclust(1-var.dist)

```

Plot the results of the clustering, for all variables that fall below the threshold, choose one from the cluster. Pick the variable that is more ecologically relevant for the questions you are asking.
```{r}
plot(var.cluster)
abline(h=0.25, lty=2, lwd=2)
```

In this example, only bio_3 and bio_4 were below the threshold. I eliminated bio_3 for the ECOSPAT.